---
title: "Comparison of Methods on a diverse Klebsiella dataset"
author: "Gerry Tonkin-Hill"
date: "`r Sys.Date()`"
output: 
  html_document:
    fig_width: 12
    fig_height: 8
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width=20, fig.height=12,
                      echo=TRUE, warning=FALSE, message=FALSE,
                      tidy=TRUE)
options(stringsAsFactors = FALSE)
```

##Load Libraries and Download Data

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
library(data.table)
library(purrr)
library(RColorBrewer)
library(aricode)
library(piggyback)
library(seqinr)
library(RWebLogo)
```

```{r, eval=FALSE}
pb_download("klebsiella_pa_files.zip", 
            repo = "gtonkinhill/panaroo_manuscript", 
            dest = "./")
unzip("klebsiella_pa_files.zip")
```

To compare the various tools on a diverse dataset we made use of 328 Klebsiella pneumoniae genomes collected from both human and animal sources across four continents. 


**Holt,K.E., Wertheim,H., Zadoks,R.N., Baker,S., Whitehouse,C.A., Dance,D., Jenney,A., Connor,T.R., Hsu,L.Y., Severin,J., et al. (2015) Genomic analysis of diversity, population structure, virulence, and antimicrobial resistance in Klebsiella pneumoniae, an urgent threat to public health. Proc. Natl. Acad. Sci. U. S. A., 112, E3574–81.**

###Run tools

Prokka was run on each assembly using an initial BLAST annotation database built for Klebsiella.

Next, each tool was run on the dataset. Panaroo was run in both strict and relaxed modes as well as being run seperately for the different species included in this dataset before combining the results using the merge tool.

```
roary -p 24 -f roary_out ./gff_files/*.gff

PIRATE -t 24 -o pirate_out -i ./gff_files/
subsample_outputs.pl -i PIRATE.gene_families.ordered.tsv -g modified_gffs/ -o ../pirate_gene_presence_absence.csv --field prev_ID

python ~/panaroo/scripts/run_panx.py -t 24 -i ./gff_files/*.gbk -o panx_out

panaroo -i ./gff_files/*.gff -o panaroo_all_out -t 24
panaroo -i ./gff_files/*.gff -o panaroo_all_sensitive -t 24 --mode relaxed
```

To merge datasets each one must be first analysed using the panaroo pipeline without cleaning up edges.

```
panaroo -i ./kpI/*.gff -o panaroo_kpI -t 24 --no_clean_edges
panaroo -i ./kpII/*.gff -o panaroo_kpII -t 24 --no_clean_edges
panaroo -i ./kpIII/*.gff -o panaroo_kpIII -t 24 --no_clean_edges

panaroo-merge -d panaroo_kpI panaroo_kpII panaroo_kpIII -o panaroo_merged -t 24
```

###Load results

```{r}
panaroo <- fread("./klebsiella_pa_files/panaroo_gene_presence_absence.csv", data.table = FALSE)
panaroo_sensitive <- fread("./klebsiella_pa_files/panaroo_sensitive_gene_presence_absence.csv", data.table = FALSE)
pirate <- fread("./klebsiella_pa_files/pirate_gene_presence_absence.csv", data.table = FALSE)
panx <- fread("./klebsiella_pa_files/panx_gene_presence_absence.csv", data.table = FALSE)
roary <- fread("./klebsiella_pa_files/roary_gene_presence_absence.csv", data.table = FALSE)
cog <- fread("./klebsiella_pa_files/cogsoft_gene_presence_absence.csv", data.table = FALSE)

pa_tables <- list(
  panaroo=panaroo[,15:ncol(panaroo)],
  panaroo_sensitive=panaroo_sensitive[,15:ncol(panaroo)],
  roary=roary[,15:ncol(roary)],
  pirate=pirate[,23:ncol(pirate)],
  panx=panx,
  cog=cog
)
```

###Distribution of Core/Accessory

```{r}
sizes_df <- imap_dfr(pa_tables, ~ tibble(
                            method=.y,
                            total=nrow(.x), 
                            core=sum(rowSums(.x!="")>(0.99*ncol(.x))), 
                            accessory=sum(rowSums(.x!="")<=(0.99*ncol(.x)))))

# correct for the way PIRATE presents duplications
sizes_df$accessory[sizes_df$method=="pirate"] <- sizes_df$accessory[sizes_df$method=="pirate"] +
  sum(apply(pa_tables[['pirate']], 1, function(r) {
    r <- r[grepl(".*;.*", r) & (!grepl(".*[()].*", r))]
    return(max(c(map_dbl(r, ~ str_count(.x, ";")), 0)))
    }))

# create a nice plot
sizes_df$method[sizes_df$method=="panaroo"] <- "Panaroo"
sizes_df$method[sizes_df$method=="panaroo_sensitive"] <- "Panaroo (sensitive)"
sizes_df$method[sizes_df$method=="panx"] <- "PanX"
sizes_df$method[sizes_df$method=="pirate"] <- "PIRATE"
sizes_df$method[sizes_df$method=="roary"] <- "Roary"
sizes_df$method[sizes_df$method=="cog"] <- "COGsoft"

plot_df <- melt(sizes_df, id.vars="method", variable.name="category", value.name="count")
plot_df$method <- factor(plot_df$method, levels = c("Panaroo","Panaroo (sensitive)", "PanX", "PIRATE", "Roary", "COGsoft"))
ggplot(plot_df, aes(x=method, y=count, fill=method)) + 
  geom_col() +
  facet_wrap(~category, ncol = 1, scales = "free_y") + 
  theme_bw(base_size = 18) +
  scale_fill_manual(values=c("#E41A1C","#FB9A99", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00"))
```

###Agreement Between Methods

As panaroo is the only method that can both remove and refind annotations we restrict this comparison to gene annotations that are present in across all methods.

```{r}
#get shared annotations
all_annotations <- unique(unlist(pa_tables[['panaroo']]))
all_annotations <- all_annotations[(all_annotations!="") & !grepl(".*refound.*", all_annotations)]
all_annotations <- unlist(map(all_annotations, ~ str_split(.x, ";")))

temp_anno <- unique(unlist(pa_tables[['panaroo_sensitive']]))
temp_anno <- temp_anno[temp_anno!=""]
all_annotations <- all_annotations[all_annotations %in% temp_anno]

temp_anno <- unique(unlist(pa_tables[['cog']]))
temp_anno <- temp_anno[temp_anno!=""]
all_annotations <- all_annotations[all_annotations %in% temp_anno]

temp_anno <- unique(unlist(pa_tables[['roary']]))
temp_anno <- temp_anno[temp_anno!=""]
all_annotations <- all_annotations[all_annotations %in% temp_anno]

temp_anno <- unique(unlist(pa_tables[['pirate']]))
temp_anno <- temp_anno[temp_anno!=""]
temp_anno <- gsub("[()]", "", unlist(str_split(unlist(temp_anno), ";|:")))
all_annotations <- all_annotations[all_annotations %in% temp_anno]

temp_anno <- unique(unlist(pa_tables[['panx']]))
temp_anno <- temp_anno[temp_anno!=""]
all_annotations <- all_annotations[all_annotations %in% temp_anno]

# determine clustering ids based on rows
clusterings <- map(pa_tables, function(pa){
  temp <- str_split(unlist(t(pa)), ";|:")
  cluster <- rep(1:nrow(pa), each=ncol(pa))
  r <- map_int(temp, length)
  
  temp <- tibble(
    gene=gsub("[()]", "", unlist(temp)),
    cluster=rep(cluster, r))
  temp <- temp[temp$gene %in% all_annotations,]
  temp <- temp[match(all_annotations, temp$gene),]
  
  return(temp$cluster)
})

# calculate Rand Index between each method
result <- matrix(NA, nrow = length(clusterings), ncol = length(clusterings),
            dimnames = list(names(clusterings), names(clusterings)))
for (i in 1:(length(clusterings)-1)){
  for (j in (i+1):length(clusterings)){
    result[i,j] <- aricode::RI(clusterings[[i]], clusterings[[j]])
    result[j,i] <- result[i,j]
  }
}

# plot the results
plotdf <- melt(result, value.name="Rand Index")
l <- unique(plotdf$Var1)
plotdf$Var1 <- factor(plotdf$Var1, levels = l)
plotdf$Var2 <- factor(plotdf$Var2, levels = l)
ggplot(data = plotdf, aes(x=Var1, y=Var2, fill=`Rand Index`)) + 
  geom_tile() +
  theme_bw(base_size = 14) +
  xlab("") + ylab("")
```


It is difficult to come up with a sensible metric for comparing methods on such a complicated dataset. One imperfect solution is to compare the number of clusters which contain genes with conflicting annotations.
We exlude genes that have been called as fragments of a larger gene as well as hypothetical proteins as these are more likely to be the result of annotation errors and are hard to sensibly compare. We then count the number of conflicting annotation in each cluster using a script included with Panaroo. It should be noted that this approach may pick up small annotation differences that aren't nesescarily clustering issues. However, overall this can give some idea whether certain methods are overclustering the genes.


First run the script

```
python ~/panaroo/scripts/count_annotation_conflicts.py -g ./panaroo_out/gene_data.csv -p ./klebsiella_pa_files/panaroo_strict_gene_presence_absence_redone.csv --method panaroo
python ~/panaroo/scripts/count_annotation_conflicts.py -g ./panaroo_out/gene_data.csv -p ./klebsiella_pa_files/pirate_gene_presence_absence.csv --method pirate
python ~/panaroo/scripts/count_annotation_conflicts.py -g ./panaroo_out/gene_data.csv -p ./klebsiella_pa_files/panx_gene_presence_absence.csv --method panx
python ~/panaroo/scripts/count_annotation_conflicts.py -g ./panaroo_out/gene_data.csv -p ./klebsiella_pa_files/roary_gene_presence_absence.csv --method roary
python ~/panaroo/scripts/count_annotation_conflicts.py -g ./panaroo_out/gene_data.csv -p ./klebsiella_pa_files/cogsoft_gene_presence_absence.csv --method cogsoft
```

Now plot the results. We copy the output by hand here.

```{r}
conflict_anno <- tibble(method=c("Panaroo","Panaroo (sensitive)", "PanX", "PIRATE", "Roary", "COGsoft"),
                        count=c(45,45,67,90,82,157))

conflict_anno$method <- factor(conflict_anno$method, levels = c("Panaroo","Panaroo (sensitive)", "PanX", "PIRATE", "Roary", "COGsoft"))
ggplot(conflict_anno, aes(x=method, y=count, fill=method)) +
  geom_col() +
  theme_bw(base_size = 18) +
  ylab("Conflicting Annotation Count") +
  scale_fill_manual(values=c("#E41A1C","#FB9A99", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00"))
```

We can also look at a couple of trickier examples to get an idea of how each method handles them.

First we take any cluster or gene that is linked to "EGEKHPPH_03354" which is most consistently annotated as 'crotonobetaine/carnitine-CoA ligase'. This is just one trickier example I selected for closer inspection.

First get a list of all genes involved

```{r, eval=FALSE}
# first get rows associated with the target
genes <- unlist(map(pa_tables, ~ unlist(.x[which(rowSums(.x=="EGEKHPPH_03354")>0),])))
genes <- unique(genes[genes!=""])

# now add genes in rows associated with those rows.
genes <- unlist(map(genes, function(g) unique(unlist(map(pa_tables, ~ unlist(.x[which(rowSums(.x==g)>0),]))))))
genes <- unique(unlist(str_split(unique(genes), "[;:]")))
genes <- unique(gsub("[()]", "", genes))

# for each gene genrate a binary outcome for being found with/without the centroid gene
is_in <- do.call(cbind, map(pa_tables, function(pa) map_lgl(genes, ~ .x %in% pa[which(rowSums(pa=="EGEKHPPH_03354")>0),])))

outtable <- cbind(genes, is_in)
outtable <- outtable[outtable[,1]!="",]
write.csv(outtable, file = "klebsiella_pa_files/EGEKHPPH_03354_pa.csv", quote = FALSE, col.names = TRUE, row.names = FALSE)
```

We use this list to generate a MSA using python, bash and MAFFT.

```{python, eval=FALSE}
genes = set(['BLIADDJL_02275','HKAMNNIE_04207','JIMHPKFH_03837','AOMCCKAJ_04206','DDAEKOIA_04666','GIBDEJGH_05261','OEAHODGD_00605','FBNNEDFL_03002','BPEFLPNN_03717','FOLCJMAC_03766','NBALKPJC_03170','MKPAIOMO_03535','ILAAFBJJ_04680','AKGJPILL_04008','DJGCGFNN_00649','HPCJGAHE_05150','JHABCCEE_05075','MCGNODNB_04349','LNFPPLKA_05107','DHJPEPJF_02985','MDCAGHGJ_00257','PDMPMOEH_04800','','KFIOGGAB_01785','JGFEPBDE_02910','BNEGIOFI_03414','AOGOPCMH_01434','GJGLCFAO_05298','CMJPCFDL_04524','ILDGAAPG_01547','ENAPNEFI_01105','HOJMNMAL_01890','AJIIOGAI_01287','OGNEABEE_00488','NKCPPPOI_02708','JMONEBNC_03164','CBCADEDL_01295','FFCOEOHO_00671','HAIFBDOH_03026','CNDDKKDK_01085','CGCLLCMO_05164','PAKMPBGM_04392','FIEOJJFB_03847','ALIHNGNC_04836','MAHDMLFK_04890','EMLJKJMF_04705','PPNCGJCJ_04346','EMNKFBDA_01170','NALFBGJE_03758','DCADFKON_04653','ACMLLLMG_01600','EICMKKAB_01899','KFGJNBPG_03255','JPPOHJAN_05116','PEFIOLCF_04113','IKCOHLNA_04900','ALFCDJGA_04938','POBLOJBP_05437','HFKIIOHA_02295','HHHPKLDM_04803','MAAKNFGI_01911','PAMKKCDO_05018','EAKINOIB_04266','EIGFAOCB_03730','GPACKGOK_04137','APNNDKAL_05194','CEMOMJPN_04915','MMPLNHOC_02699','PKCIBKII_03201','PACKLOLK_04787','HBEIADDC_02948','ODLCBFAL_02402','JPLOIOGN_01029','NNCHLEOD_05112','INFPFCGE_01342','GNGDDAIF_00851','ECHNOMEK_01597','CAGJNEIA_04680','NJAAJKJA_05348','MBLCGNJA_03933','DHGKKBIJ_04886','DCMCHNCH_00666','KCFKDDPO_04739','HBNMADGG_04953','PIMKEAGC_01627','ILDMPJJF_04882','DLHFOGND_04266','AHIKKBJF_05064','JGOJCHHP_05000','BNNFCJDG_03596','FEIFONLG_04499','DLICPGHM_05032','PPHDDNLI_00531','GNMAMBFN_05134','MNNOIGOP_02518','PEDJOJGP_01350','CNIEKNAI_03812','AKKCFLDB_03574','FKLLCIKP_00215','MHPGDIGD_04353','FGIHKDBE_01606','MMIGMJEO_02924','PKCMIDDC_04281','LLOPFMIA_00992','MOLGLLGN_01280','CMALKGEE_02649','JMPGAFMA_02144','LOJEIGOF_04458','GKFAMBLJ_03798','EHLIBNLI_03103','BLHJFOEP_04117','HFEJPIEE_01716','HGJCJHNL_04238','EPEODNLN_03924','CGAIOHOI_02687','JNPIPOAJ_00200','MHGIHCEB_03832','FFHEGANC_03650','MBPCNGAM_02835','PDJEFONO_04954','IIMJLGKH_00810','LHFOCOKJ_02630','ODAAEODD_04048','OHFDNAHM_02623','LMLMECGJ_04519','FGGOHEAC_00218','FBFHPNLF_02254','PGCIPBHI_04286','FOEGGEAC_03039','DNHMAGKA_04934','MCAFBMFF_05003','KDJNOHOH_04560','KBMHKIEM_04888','LDPGPIJA_01908','LBJBIDEJ_04927','HFKIIAGP_04614','KMJFAHHH_03317','DLBLDKIJ_03683','HMJNIDCC_02663','LEIJEFKP_01593','MFNKMPAM_02026','BAMHCAME_02962','CAFAJEPH_03897','AIDFMLNP_03864','ECJFIBNB_03311','JGJMBNEK_04178','GCFPCKHP_03091','NDPMMGBI_04757','CDCEHFGC_03781','MLKBEPLK_01220','CJFFOHAO_03781','GDFNPJKD_03417','FENLIFPN_03787','GPHMKBKC_03768','BEHKDFBL_04851','LNPOPIPF_03163','IFPDIJNH_04138','IGOGDGOA_03653','BFFFBEKE_02897','MIFHILDP_00377','PDEDMGPD_03953','IIKLHBBD_03485','HEHNFIEG_04095','JODHPGGM_03160','KIKMJPAP_02006','CHLEAICA_03543','MCMNCFNH_04974','IPGNFAMM_03974','OJFONEFE_02852','CPFPLKLM_01089','ILDBFKKF_01448','MLJDNLLD_01350','IMGDPOBN_00835','EHMCBADO_04735','AEHBFAPC_04957','NBAGIKNE_04786','FIMDDPAE_04972','HHCNOMOF_03453','FOILALFG_04914','CBNBDEKD_03454','EGGCBBJP_03449','CHGKJLKP_02994','EJJBJLNA_00964','IDPGEAOI_00578','HAJJLEIN_04320','MFFEMMHI_01471','GLGBDILJ_03422','LKPEOILP_05580','JPDAEMJH_05434','HFDNAOLN_05064','HJOBFDCO_04546','HJOBFDCO_04545','FEMCEJGN_05499','NFPLLLIB_03027','BHJEDKBF_05297','MIJDJFOI_04307','DGEBHCBB_05360','GHBFKOKL_05295','IJODNBMD_05521','DJPFCFMD_04307','HMIKBBDF_05416','BGLJCGEP_04489','HNIOAKBO_00374','FMEGIGHO_04799','FMEGIGHO_00614','NMPKMPNM_02470','GFBGDJHB_05091','BDMAJJCI_05301','NLKOGHBO_04924','BFDFNMAI_04828','BFDFNMAI_04827','AMPEBOID_05153','HGOIIGDC_04518','GDDIMEOF_04945','AJBMBHGB_04787','IJAAPAFL_04202','LPFMOLNC_05228','KKAGLBML_05349','AIINIPAH_03895','GMPGEAAC_04911','NKFOMIGP_05323','LLOJMOCI_05214','FAPHKOEP_04905','DPNFLAPC_04405','EKENHPDP_04769','NNHDGMCP_00057','OHCKPMKE_04798','FAGJOMKD_05011','LNOJILEH_05067','LNOJILEH_05068','NEDPJPIG_05072','KFMJDIAC_05124','GLCLDGNF_04996','CINLLEGC_02776','OEFPCOAF_05329','OEFPCOAF_02242','BOPIEMCD_04863','PPJHPCJK_05064','GLBFKCFA_03563','GLBFKCFA_03564','BIDGGFEG_04444','BIDGGFEG_04445','NIDGGJLL_05039','NIDGGJLL_05040','EEDMPDLB_02829','FEHCLKIG_01696','DNGHDJHL_04660','KEAACCIK_05314','LAFDFEKJ_02089','KHNPOKNE_04964','BFIPIBFD_02588','HOFOBAGB_04436','FCIECHGP_04770','FCIECHGP_04771','PMEGGNDC_01869','LEJHAAAP_04663','PCNBLMKD_04960','DEHOHPIP_04484','LBHCMNPC_02244','KOAILEED_01359','GMMKEPEF_03084','GMMKEPEF_03085','NJHLNCFD_01255','EBJBKJPC_03299','OBOFFKJP_01760','OBOFFKJP_01759','GANEBHNN_02734','AEJAMGBJ_04585','FMHOMKLA_04888','JCIHFHBB_02821','PDHINEDK_03394','FAMLPOHJ_04036','MOOPCBGB_00313','FPOPEOBJ_02073','HINBENMJ_03926','OCADADBM_01034','LEOFLMDA_00193','DIBGMLIL_04333','IDKJIBFA_04853','ABKIGFEM_03363','IENNIAEH_04216','KCCKIDPK_04442','MCCBNMDM_02482','PGLCGBHE_00971','FIKNJMIL_00602','ELLIMNIF_02643','DMAHHBKC_00566','BEJEIACC_03270','KICOGNBF_03550','HJCAAAAL_02524','PCKJBFPB_03715','MHBMGKGF_03397','GACEEEAN_02594','EGEKHPPH_03354','CLNFAHOL_03348','JAAJOMDF_03696','GPLELKNM_03741','IEKEKMED_03842','EFJGEALO_03337','HHGKAPAN_02583','KNKLCFIP_02655','GMINOEHK_03762','KPHOFCDF_00065','KPOIAHFD_00064','OLPAEAFJ_02724','NKMKOHOB_00375','LFLMDOGK_04815','HBGICMEK_04560','HBGICMEK_04561','LFDFGGNM_03648','ADCANDME_00670','IGCBFHDK_04836','FOHJBLJH_03649','KAILDODD_00001','MJLPJJDP_05240','JLDIOFAI_05261','DBDDMCCB_05046','LNBGOKMO_04416','MGBABOJF_05274','NFPLLLIB_01325','BMANODNA_04718','BHJEDKBF_05260','MIJDJFOI_01188','BPAPPDLG_05054','BDMAJJCI_05406','NLKOGHBO_05218','HGOIIGDC_04517','AJBMBHGB_05446','LGLBPKJJ_04884','LIDJBJFH_03636','BPKAMOIJ_04912','HHBHKLFP_05372','HHBHKLFP_05385','GLCLDGNF_05211','GLCLDGNF_05212','BOPIEMCD_05334','KEAACCIK_05315','OHAGMEMK_04936','OHAGMEMK_05327','DEHOHPIP_05101','JNJKFODN_05144','DDBMEDAI_05599','209_refound_18115','287_refound_27274','0_173_refound_3835','0_246_refound_7373'])

with open("gene_data.csv", 'r') as infile, \
  open("EGEKHPPH_03354.fasta", 'w') as outfile:
  next(infile)
  for line in infile:
    line=line.split(",")
    if line[3] in genes:
      outfile.write(">" + line[3] + "\n" +
        line[4] + "\n")
```


```
mafft EGEKHPPH_03354.fasta > EGEKHPPH_03354_MSA.fasta
```

```{r}
msa <-  do.call(rbind, seqinr::read.fasta("./klebsiella_pa_files/EGEKHPPH_03354_MSA.fasta", 
                                          seqtype = "AA", as.string = TRUE))
method_pa <- fread("./klebsiella_pa_files/EGEKHPPH_03354_pa.csv", data.table = FALSE)
method_pa <- method_pa[match(rownames(msa), method_pa$genes),]

method_pa[(method_pa$panaroo==TRUE) & (method_pa$pirate==FALSE),]
method_pa[(method_pa$panaroo==FALSE) & (method_pa$pirate==TRUE),]
msa[method_pa$panaroo!=method_pa$pirate,]

RWebLogo::weblogo(msa[(method_pa$panaroo==TRUE) & (method_pa$pirate==TRUE),], composition="none", scale.width=FALSE, stacks.per.line = 1000, size = "small",
                  units="probability", file.out = "./Figures/EGEKHPPH_03354_panarooT_pirateT.pdf", format = "pdf")
RWebLogo::weblogo(msa[(method_pa$panaroo==FALSE) & (method_pa$pirate==TRUE),], composition="none", scale.width=FALSE, stacks.per.line = 1000, size = "small",
                  units="probability", file.out = "./Figures/EGEKHPPH_03354_panarooF_pirateT.pdf", format = "pdf")
RWebLogo::weblogo(msa[(method_pa$panaroo==TRUE) & (method_pa$pirate==FALSE),], composition="none", scale.width=FALSE, stacks.per.line = 1000, size = "small",
                  units="probability", file.out = "./Figures/EGEKHPPH_03354_panarooT_pirateF.pdf", format = "pdf")

RWebLogo::weblogo(msa[(method_pa$panaroo==TRUE) & (method_pa$panx==TRUE),], composition="none", scale.width=FALSE, stacks.per.line = 1000, size = "small",
                  units="probability", file.out = "./Figures/EGEKHPPH_03354_panarooT_panxT.pdf", format = "pdf")
RWebLogo::weblogo(msa[(method_pa$panaroo==FALSE) & (method_pa$panx==TRUE),], composition="none", scale.width=FALSE, stacks.per.line = 1000, size = "small",
                  units="probability", file.out = "./Figures/EGEKHPPH_03354_panarooF_panxT.pdf", format = "pdf")
RWebLogo::weblogo(msa[(method_pa$panaroo==TRUE) & (method_pa$panx==FALSE),], composition="none", scale.width=FALSE, stacks.per.line = 1000, size = "small",
                  units="probability", file.out = "./Figures/EGEKHPPH_03354_panarooT_panxF.pdf", format = "pdf")

RWebLogo::weblogo(msa[(method_pa$panaroo==TRUE) & (method_pa$roary==TRUE),], composition="none", scale.width=FALSE, stacks.per.line = 1000, size = "small",
                  units="probability", file.out = "./Figures/EGEKHPPH_03354_panarooT_roaryT.pdf", format = "pdf")
RWebLogo::weblogo(msa[(method_pa$panaroo==FALSE) & (method_pa$roary==TRUE),], composition="none", scale.width=FALSE, stacks.per.line = 1000, size = "small",
                  units="probability", file.out = "./Figures/EGEKHPPH_03354_panarooF_roaryT.pdf", format = "pdf")
RWebLogo::weblogo(msa[(method_pa$panaroo==TRUE) & (method_pa$roary==FALSE),], composition="none", scale.width=FALSE, stacks.per.line = 1000, size = "small",
                  units="probability", file.out = "./Figures/EGEKHPPH_03354_panarooT_roaryF.pdf", format = "pdf")

# RWebLogo::weblogo(msa[method_pa$panaroo==method_pa$cogsoft,], composition="none", scale.width=FALSE, stacks.per.line = 1000, size = "small",
#                   units="probability", file.out = "./Figures/EGEKHPPH_03354_panarooT_cogT.pdf", format = "pdf")
# RWebLogo::weblogo(msa[(method_pa$panaroo==FALSE) & (method_pa$cogsoft==TRUE),], composition="none", scale.width=FALSE, stacks.per.line = 1000, size = "small",
#                   units="probability", file.out = "./Figures/EGEKHPPH_03354_panarooF_cogT.pdf", format = "pdf")
# RWebLogo::weblogo(msa[(method_pa$panaroo==TRUE) & (method_pa$cogsoft==FALSE),], composition="none", scale.width=FALSE, stacks.per.line = 1000, size = "small",
#                   units="probability", file.out = "./Figures/EGEKHPPH_03354_panarooT_cogF.pdf", format = "pdf")
```

We can also look at "PDHINEDK_04333" which is less conserved and most consistently annotated as a 'transcriptional regulator'.

```{r, eval=FALSE}
# first get rows associated with the target
genes <- unlist(map(pa_tables, ~ unlist(.x[which(rowSums(.x=="PDHINEDK_04333")>0),])))
genes <- unique(genes[genes!=""])

# now add genes in rows associated with those rows.
genes <- unlist(map(genes, function(g) unique(unlist(map(pa_tables, ~ unlist(.x[which(rowSums(.x==g)>0),]))))))
genes <- unique(unlist(str_split(unique(genes), "[;:]")))
genes <- unique(gsub("[()]", "", genes))

# for each gene genrate a binary outcome for being found with/without the centroid gene
is_in <- do.call(cbind, map(pa_tables, function(pa) map_lgl(genes, ~ .x %in% pa[which(rowSums(pa=="PDHINEDK_04333")>0),])))

outtable <- cbind(genes, is_in)
outtable <- outtable[outtable[,1]!="",]
write.csv(outtable, file = "klebsiella_pa_files/PDHINEDK_04333_pa.csv", quote = FALSE, row.names = FALSE)
```

```{python, eval=FALSE}
genes = set(['FBNNEDFL_04560','AKGJPILL_04330','MDCAGHGJ_04733','KFIOGGAB_03125','JGFEPBDE_02632','CMJPCFDL_01003','AJIIOGAI_00991','NKCPPPOI_00877','CNDDKKDK_02962','PAKMPBGM_00831','EMLJKJMF_04052','PEFIOLCF_00949','JLDIOFAI_04140','HHHPKLDM_01101','GPACKGOK_04048','APNNDKAL_02393','CEMOMJPN_03598','MMPLNHOC_03355','PKCIBKII_03483','JPLOIOGN_00025','GNGDDAIF_00461','ECHNOMEK_00291','CAGJNEIA_02194','MBLCGNJA_04625','BNNFCJDG_04761','GNMAMBFN_04771','PEDJOJGP_04644','MMIGMJEO_04043','PKCMIDDC_01051','LLOPFMIA_04709','GKFAMBLJ_04201','LHFOCOKJ_03172','MCAFBMFF_04546','ECJFIBNB_00755','CHLEAICA_01884','EHMCBADO_04650','FOILALFG_00032','LNBGOKMO_00840','LKPEOILP_03269','BHJEDKBF_01348','MIJDJFOI_02327','DGEBHCBB_03804','HNIOAKBO_00932','GMPGEAAC_00059','NKFOMIGP_04029','FAPHKOEP_03377','DPNFLAPC_02528','EKENHPDP_03837','NEDPJPIG_03454','CINLLEGC_03368','OEFPCOAF_03439','PPJHPCJK_04406','GLBFKCFA_04023','DNGHDJHL_02871','LAFDFEKJ_03357','NJHLNCFD_01136','GANEBHNN_03901','AEJAMGBJ_03399','FMHOMKLA_03599','PDHINEDK_04333','DHJPEPJF_03747','MDCAGHGJ_04635','PDMPMOEH_03367','KFIOGGAB_04582','CNDDKKDK_04492','ALIHNGNC_03772','MAHDMLFK_03598','PPNCGJCJ_03750','GPACKGOK_03337','APNNDKAL_02646','INFPFCGE_00087','NJAAJKJA_04086','MBLCGNJA_03828','DHGKKBIJ_00004','DLHFOGND_03200','JGOJCHHP_01778','BNNFCJDG_03807','FEIFONLG_02516','CNIEKNAI_00328','CNIEKNAI_00332','MOLGLLGN_03496','OHFDNAHM_04698','JGJMBNEK_00252','JGJMBNEK_00256','CDCEHFGC_00697','CDCEHFGC_00701','CJFFOHAO_01579','CJFFOHAO_01583','FENLIFPN_01287','FENLIFPN_01291','GPHMKBKC_01031','GPHMKBKC_01035','OJFONEFE_00120','OJFONEFE_00124','MLJDNLLD_03867','IMGDPOBN_05085','NBAGIKNE_01437','NBAGIKNE_01441','FIMDDPAE_03061','MFFEMMHI_02606','MFFEMMHI_02610','HJOBFDCO_00391','HJOBFDCO_00396','HGOIIGDC_02117','KKAGLBML_00063','KKAGLBML_00067','AIINIPAH_01698','NNHDGMCP_01527','KFMJDIAC_00335','CINLLEGC_02904','OEFPCOAF_03047','PPJHPCJK_02419','GLBFKCFA_03447','BIDGGFEG_00080','NIDGGJLL_00878','LAFDFEKJ_03213','KHNPOKNE_00183','PMEGGNDC_01230','KOAILEED_01900','OBOFFKJP_00714','GANEBHNN_00560','FPOPEOBJ_03531','DDBMEDAI_00284','DDBMEDAI_00288','FIKNJMIL_00988','FIKNJMIL_00992','ELLIMNIF_05254','ELLIMNIF_05258','DMAHHBKC_05270','DMAHHBKC_05274','BEJEIACC_05243','BEJEIACC_05247','KICOGNBF_04851','KICOGNBF_04855','HJCAAAAL_05283','HJCAAAAL_05287','PCKJBFPB_01381','PCKJBFPB_01385','MHBMGKGF_05270','MHBMGKGF_05274','GACEEEAN_00638','GACEEEAN_00642','EGEKHPPH_02915','EGEKHPPH_02919','CLNFAHOL_01088','CLNFAHOL_01092','JAAJOMDF_01086','JAAJOMDF_01090','GPLELKNM_01396','GPLELKNM_01400','IEKEKMED_05248','IEKEKMED_05252','EFJGEALO_05294','EFJGEALO_05298','HHGKAPAN_00749','HHGKAPAN_00753','KNKLCFIP_00712','KNKLCFIP_00716','GMINOEHK_05085','GMINOEHK_05089','LFLMDOGK_04163','HBGICMEK_01221','HBGICMEK_02155','LFDFGGNM_00804','LFDFGGNM_00808','ADCANDME_01885','IGCBFHDK_03794','FOHJBLJH_01221','FOHJBLJH_01225'])

with open("gene_data.csv", 'r') as infile, \
  open("PDHINEDK_04333.fasta", 'w') as outfile:
  next(infile)
  for line in infile:
    line=line.split(",")
    if line[3] in genes:
      outfile.write(">" + line[3] + "\n" +
        line[4] + "\n")
```

```{r}
msa <-  do.call(rbind, seqinr::read.fasta("./klebsiella_pa_files/PDHINEDK_04333_MSA.fasta", 
                                          seqtype = "AA", as.string = TRUE))
method_pa <- fread("./klebsiella_pa_files/PDHINEDK_04333_pa.csv", data.table = FALSE)
method_pa <- method_pa[match(rownames(msa), method_pa$genes),]

method_pa[(method_pa$panaroo==TRUE) & (method_pa$pirate==FALSE),]
method_pa[(method_pa$panaroo==FALSE) & (method_pa$pirate==TRUE),]
msa[method_pa$panaroo!=method_pa$pirate,]

RWebLogo::weblogo(msa[(method_pa$panaroo==TRUE) & (method_pa$pirate==TRUE),], composition="none", scale.width=FALSE, stacks.per.line = 1000, size = "small",
                  units="probability", file.out = "./Figures/PDHINEDK_04333_panarooT_pirateT.pdf", format = "pdf")
RWebLogo::weblogo(msa[(method_pa$panaroo==FALSE) & (method_pa$pirate==TRUE),], composition="none", scale.width=FALSE, stacks.per.line = 1000, size = "small",
                  units="probability", file.out = "./Figures/PDHINEDK_04333_panarooF_pirateT.pdf", format = "pdf")
RWebLogo::weblogo(msa[(method_pa$panaroo==TRUE) & (method_pa$pirate==FALSE),], composition="none", scale.width=FALSE, stacks.per.line = 1000, size = "small",
                  units="probability", file.out = "./Figures/PDHINEDK_04333_panarooT_pirateF.pdf", format = "pdf")

RWebLogo::weblogo(msa[(method_pa$panaroo==TRUE) & (method_pa$panx==TRUE),], composition="none", scale.width=FALSE, stacks.per.line = 1000, size = "small",
                  units="probability", file.out = "./Figures/PDHINEDK_04333_panarooT_panxT.pdf", format = "pdf")
RWebLogo::weblogo(msa[(method_pa$panaroo==FALSE) & (method_pa$panx==TRUE),], composition="none", scale.width=FALSE, stacks.per.line = 1000, size = "small",
                  units="probability", file.out = "./Figures/PDHINEDK_04333_panarooF_panxT.pdf", format = "pdf")
RWebLogo::weblogo(msa[(method_pa$panaroo==TRUE) & (method_pa$panx==FALSE),], composition="none", scale.width=FALSE, stacks.per.line = 1000, size = "small",
                  units="probability", file.out = "./Figures/PDHINEDK_04333_panarooT_panxF.pdf", format = "pdf")

RWebLogo::weblogo(msa[(method_pa$panaroo==TRUE) & (method_pa$roary==TRUE),], composition="none", scale.width=FALSE, stacks.per.line = 1000, size = "small",
                  units="probability", file.out = "./Figures/PDHINEDK_04333_panarooT_roaryT.pdf", format = "pdf")
RWebLogo::weblogo(msa[(method_pa$panaroo==FALSE) & (method_pa$roary==TRUE),], composition="none", scale.width=FALSE, stacks.per.line = 1000, size = "small",
                  units="probability", file.out = "./Figures/PDHINEDK_04333_panarooF_roaryT.pdf", format = "pdf")
RWebLogo::weblogo(msa[(method_pa$panaroo==TRUE) & (method_pa$roary==FALSE),], composition="none", scale.width=FALSE, stacks.per.line = 1000, size = "small",
                  units="probability", file.out = "./Figures/PDHINEDK_04333_panarooT_roaryF.pdf", format = "pdf")

# RWebLogo::weblogo(msa[method_pa$panaroo==method_pa$cogsoft,], composition="none", scale.width=FALSE, stacks.per.line = 1000, size = "small",
#                   units="probability", file.out = "./Figures/EGEKHPPH_03354_panarooT_cogT.pdf", format = "pdf")
# RWebLogo::weblogo(msa[(method_pa$panaroo==FALSE) & (method_pa$cogsoft==TRUE),], composition="none", scale.width=FALSE, stacks.per.line = 1000, size = "small",
#                   units="probability", file.out = "./Figures/EGEKHPPH_03354_panarooF_cogT.pdf", format = "pdf")
# RWebLogo::weblogo(msa[(method_pa$panaroo==TRUE) & (method_pa$cogsoft==FALSE),], composition="none", scale.width=FALSE, stacks.per.line = 1000, size = "small",
#                   units="probability", file.out = "./Figures/EGEKHPPH_03354_panarooT_cogF.pdf", format = "pdf")
```






We can investigate differences in the core genome calls between the methods.

```{r, eval=FALSE}
core_panaroo <- pa_tables[["panaroo"]][rowSums(pa_tables[["panaroo"]]!="")>(0.95*ncol(pa_tables[["panaroo"]])),]
core_pirate <-  pa_tables[["panx"]][rowSums(pa_tables[["panx"]]!="")>(0.95*ncol(pa_tables[["panx"]])),]

all(colnames(core_panaroo) %in% colnames(core_pirate))
all(colnames(core_pirate) %in% colnames(core_panaroo))


core_pirate <- core_pirate[,match(colnames(core_panaroo), colnames(core_pirate))]

pa_tables[["pirate"]]  <- pa_tables[["pirate"]][,match(colnames(core_panaroo), colnames(pa_tables[["pirate"]]))]
all(colnames(core_panaroo) == colnames(core_pirate))

core_pirate[core_pirate==""] <- NA
core_panaroo[core_panaroo==""] <- NA

hits <- apply(core_pirate, 1, function(r){
  n_matches <- rowSums(t(t(pa_tables[["panaroo"]])==r), na.rm = TRUE)
  return(which(n_matches>0))
})

core_panaroo_index <- which(rowSums(pa_tables[["panaroo"]]!="")>(0.95*ncol(pa_tables[["panaroo"]])))
diff <- map(hits, ~ all(.x %in% core_panaroo_index))
diff <- hits[!unlist(diff)]

bad <- imap_lgl(diff, ~ (length(.x)==1) & sum(core_pirate[as.numeric(.y),]!=pa_tables[["panaroo"]][.x[[1]],], na.rm = TRUE)>5)
diff[bad]


length(diff)
core_pirate[is.na(core_pirate)] <- ""
pirate[3581,]
sum(core_pirate[158,]==pa_tables[["panaroo"]][5154,], na.rm = TRUE)
sum(core_pirate[158,]!=pa_tables[["panaroo"]][5154,], na.rm = TRUE)

d <- apply(pa_tables[["panaroo"]], 1, function(r) sum(r==core_pirate[158,], na.rm = TRUE))
which(d>0)

pa_tables[["panaroo"]][5154,]

rbind(core_pirate[158,], pa_tables[["panaroo"]][5154,])[,core_pirate[158,]!=pa_tables[["panaroo"]][5154,]]
rbind(core_pirate[158,], pa_tables[["panaroo"]][5154,])[,core_pirate[158,]==pa_tables[["panaroo"]][5154,]]
which(rowSums(t(t(pa_tables[["pirate"]])==core_pirate[3693,]), na.rm = TRUE)>0)
d <- apply(pa_tables[["pirate"]], 1, function(r) sum(r==core_pirate[2618,], na.rm = TRUE))

which(rowSums(pa_tables[["panaroo"]]=="KAILDODD_04989")>0)
which(apply(pa_tables[["panaroo"]], 1, function(x) sum(grepl(".*KAILDODD_04989.*", x)))>0)

pa_tables[["panaroo"]][1372,]

pa_tables[["panaroo"]][1372,grepl(".*KAILDODD_04989.*", pa_tables[["panaroo"]][1372,])]

low <- which(pirate$threshold==70)
low <- low[100:length(low)]
low[1:70]
pirate[5154,]

all_pirate <- pirate[]

rbind(pa_tables[["pirate"]][5154,],
      pa_tables[["panaroo"]][158])[,pa_tables[["pirate"]][5154,]!=pa_tables[["panaroo"]][158,]]
rbind(pa_tables[["pirate"]][5154,],
      pa_tables[["panaroo"]][158,])[,pa_tables[["pirate"]][5154,]==pa_tables[["panaroo"]][158,]]



d <- apply(pa_tables[["panaroo"]], 1, function(r) sum(any(core_pirate[2618,], r), na.rm = TRUE))

d <- apply(pa_tables[["panaroo_sensitive"]], 1, function(r) sum(r==core_pirate[2618,], na.rm = TRUE))



paste(pa_tables[["pirate"]][3369,], collapse = " ")
pirate[3369,]
pa_tables[["panaroo"]][5747,]

```

```{r}

```






