---
title: "Analysis of Gene Presence/Absence in the Pneumoccocal Global Sequenceing Project"
author: "Gerry Tonkin-Hill"
date: "`r Sys.Date()`"
output: 
  html_document:
    fig_width: 12
    fig_height: 8
editor_options: 
  chunk_output_type: console
bibliography: references.bib
nocite: | 
  @Gladstone2019-rz
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width=20, fig.height=12,
                      echo=TRUE, warning=FALSE, message=FALSE,
                      tidy=TRUE)
options(stringsAsFactors = FALSE)
```

##Load Libraries and Download Data

```{r, warning=FALSE, message=FALSE}
library(data.table)
library(tidyverse)
library(RColorBrewer)
library(piggyback)
```

```{r, eval=FALSE}
pb_download("pneumo_gps_data.zip", 
            repo = "gtonkinhill/panaroo_manuscript", 
            dest = "./")
unzip("pneumo_gps_data.zip")
```

To demonstrate Panaroo's ability to scale to large datasets and to infer pangenome evolutionary paramters we investigated a subsection of the large Global Pneumococcal Sequenceing (GPS) project.

**Gladstone,R.A., Lo,S.W., Lees,J.A., Croucher,N.J., van Tonder,A.J., Corander,J., Page,A.J., Marttinen,P., Bentley,L.J., Ochoa,T.J., et al. (2019) International genomic definition of pneumococcal lineages, to contextualise disease, antibiotic resistance and vaccine impact. EBioMedicine, 43, 338–346.**

We only considered isolates from the major sequencing clusters (GPSC's) for which a reliable dated tree could be inferred. The dated phylogenies were obtained from **Gladstone,R.A., et al. (2019) in prep**.

##Run Panaroo

Panaroo was first run seperately on each of the GPSCs both in the strict and sensitive modes.

```
for folder in *
do

cd $folder

mkdir panaroo_out
panaroo -i *.gff -o panaroo_out -t 10 --verbose

mkdir panaroo_sensitive_out
panaroo -i *.gff -o panaroo_out -t 10 --verbose --mode relaxed

cd ..

done
```

The results of these analyses was then merged using the panaroo-merge command

```
mkdir panaroo_merged_out
panaroo-merge -d [0-9]* -o panaroo_merged_out -t 24 --verbose
```

We next used one of the Panaroo post processing functions to infer the paramters of the Infinitely Many Genes (IMG) model as described in @Collins2012-yf and @Baumdicker2012-wj.

```
for folder in [0-9]*
do
cd ${folder}/panaroo_out
mkdir img_model
cp ../../bactdater_outputs/GPSC${folder}_BD.tre ./
python ~/panaroo/panaroo-estimate-img.py --pa gene_presence_absence.Rtab -o img_model/ --tree GPSC${folder}_BD.tre -D 2
cd ../..
done
```

We also inferred the parameters of the Finitely Many Genes model as described in [@Zamani-Dahaj2016-gi]

```
for folder in [0-9]*
do
cd ${folder}/panaroo_out
mkdir fmg_model
python ~/panaroo/panaroo-estimate-fmg.py --model FMG --tree GPSC${folder}_BD.tre --pa gene_presence_absence_renamed.Rtab -o fmg_model/${folder}_fmg_results.txt -t 5 --nboot 100 --verbose
cd ../..
done
```

After moving the resulting files into a single directory we can now load the results.

```{r}
# get a list of the output files
fmg_param_files <- Sys.glob("./pneumo_gps_data/fmg_estimates/*.txt")
# img_size_files <- Sys.glob("./pneumo_gps_data/img_estimates/*size*")
# img_param_files <- Sys.glob("./pneumo_gps_data/img_estimates/*param*")
binary_pa_files <- Sys.glob("./pneumo_gps_data/binary_pa_matrices/*_strict_gene_presence_absence.Rtab")

# load the data in a useful format
fmg_params <- map_dfr(fmg_param_files, function(x){
                  tbl <- fread(x, 
                          data.table = FALSE, 
                          skip = 4, 
                          col.names = c("param", "estimate", "2.5CI", "97.5CI"))
                  tbl$GPSC <- gsub("_.*", "", gsub(".*/", "", x))
                  tbl$method <-  gsub(".*_", "", gsub("_fmg.*", "", x))
                  return(tbl)
                  })

#load cluster sizes
cluster_sizes <- map_dfr(binary_pa_files, function(x){
                  m <- fread(x, 
                          data.table = FALSE)
                  tbl <- data.frame(GPSC = gsub("_.*", "", gsub(".*/", "", x)),
                                    size = ncol(m) -1, 
                                    stringsAsFactors = FALSE)
                  return(tbl)
                  })
cluster_sizes <- melt(cluster_sizes)
cluster_sizes <- tibble(param=cluster_sizes$variable,
                            estimate=as.numeric(cluster_sizes$variable),
                            `2.5CI`=NA,
                            `97.5CI`=NA,
                            GPSC=as.numeric(cluster_sizes$GPSC),
                            method=rep("strict", nrow(cluster_sizes)))

# img_params <- map_dfr(img_param_files, function(x){
#   tbl <- fread(x,
#                data.table = FALSE,
#                header = TRUE)
#   tbl$GPSC <- gsub("_.*", "", gsub(".*/", "", x))
#   tbl$method <-  gsub(".*_", "", gsub("_infered.*", "", x))
#   return(tbl)
# })
# 
# img_sizes <- map_dfr(img_size_files, function(x){
#   tbl <- fread(x,
#                data.table = FALSE,
#                header = TRUE)
#   tbl$GPSC <- gsub("_.*", "", gsub(".*/", "", x))
#   tbl$method <-  gsub(".*_", "", gsub("_core.*", "", x))
#   return(tbl)
# })
```

We can also load some additional supplementary information from @Gladstone2019-rz.

```{r}
odds_ratio <- fread("./pneumo_gps_data/GPSC_OR.csv", data.table = FALSE)
sig_odds <- tibble(GPSC=odds_ratio$GPSC, log_odds=log(odds_ratio$logOR), issig=odds_ratio$pvalue<0.05)
odds_ratio <- tibble(GPSC=rep(odds_ratio$GPSC, 2),
                     param="odds_ratio",
                     estimate=rep(odds_ratio$logOR,2),
                     `2.5CI`=NA, `97.5CI`=NA,
                     method=rep(c("sensitive", "strict"), each=nrow(odds_ratio)))
recombination <- fread("./pneumo_gps_data/GPSC_recombination_T11.csv", data.table = FALSE)
recombination <- recombination[!is.na(recombination$GPSC),]
recombination <- tibble(GPSC=rep(recombination$GPSC, 2),
                     param="r/m",
                     estimate=rep(recombination$`r/m`,2),
                     `2.5CI`=NA, `97.5CI`=NA,
                     method=rep(c("sensitive", "strict"), each=nrow(recombination)))

# cluster_sizes <- img_sizes %>% group_by(GPSC) %>%
#   summarise(count=n())
# cluster_sizes <- tibble(GPSC=rep(cluster_sizes$GPSC, 2),
#                      param="cluster size",
#                      estimate=rep(cluster_sizes$count,2),
#                      `2.5CI`=NA, `97.5CI`=NA,
#                      method=rep(c("sensitive", "strict"), each=nrow(cluster_sizes)))

diversity <- fread("./pneumo_gps_data/GPSC_diversity_T8.csv", data.table = FALSE)
diversity <- tibble(GPSC=rep(diversity$GPSC, 2),
                     param="country diversity",
                     estimate=rep(diversity$`% PCV13`,2),
                     `2.5CI`=NA, `97.5CI`=NA,
                     method=rep(c("sensitive", "strict"), each=nrow(diversity)))

other_data <- fread("./pneumo_gps_data/GPSC_summary_T1.csv", data.table = FALSE)
n_serotypes <- other_data %>% group_by(GPSC) %>%
  summarise(n_sero=length(unique(In_Silico_Serotype)))
n_serotypes <- tibble(GPSC=rep(n_serotypes$GPSC, 2),
                     param="num sero",
                     estimate=rep(n_serotypes$n_sero,2),
                     `2.5CI`=NA, `97.5CI`=NA,
                     method=rep(c("sensitive", "strict"), each=nrow(n_serotypes)))
```

Merge the data together ready for plotting.

```{r}
fmg_params <- rbind(fmg_params, cluster_sizes)
fmg_params <- rbind(fmg_params, recombination[recombination$GPSC %in% fmg_params$GPSC,])
fmg_params <- rbind(fmg_params, odds_ratio[odds_ratio$GPSC %in% fmg_params$GPSC,])
fmg_params <- rbind(fmg_params, n_serotypes[n_serotypes$GPSC %in% fmg_params$GPSC,])
fmg_params <- rbind(fmg_params, diversity[diversity$GPSC %in% fmg_params$GPSC,])
```

Plot the estimated parameters for the Finitely Many Genes model along with some metadata for each GPSC

```{r}
tmp <- fmg_params[fmg_params$param=="M" & fmg_params$method=="strict",]
tmp <- tmp$GPSC[order(-tmp$estimate)]
tmp <- c(tmp, unique(fmg_params$GPSC[!(fmg_params$GPSC %in% tmp)]))
fmg_params$GPSC <- factor(fmg_params$GPSC, levels = tmp)

ggplot(fmg_params[fmg_params$method=="strict",], aes(x=GPSC, y=estimate)) + 
  geom_col() +
  facet_wrap(~ param, ncol = 1, scales = "free_y") +
  theme_bw(base_size = 14)
```

From this plot it looks like there could be a weak association between genome gain/loss rate and the odds ratio of the lineage. It also looks like there might be a weak association between the number of seroptypes, the recombination rate and the estiamted size of the pangenome pool (M). We can perform a simple 

```{r}
fmg_params$param[fmg_params$param=="odds_ratio"] <- "odds ratio"
fmg_params$param[fmg_params$param=="num sero"] <- "unique serotypes"
fmg_data <- dcast(fmg_params, method+GPSC ~ param, value.var="estimate")
fmg_data <- fmg_data[fmg_data$method=="strict",]

cor(fmg_data$M, fmg_data$`unique serotypes`, method = "spearman")
cor.test(fmg_data$M, fmg_data$`unique serotypes`, method = "spearman")

fmg_data <- fmg_data[!is.na(fmg_data$`r/m`),]
cor(fmg_data$M, fmg_data$`r/m`, method = "spearman")
cor.test(fmg_data$M, fmg_data$`r/m`, method = "spearman")
```


```{r}
fmg_data <- dcast(fmg_params, method+GPSC ~ param, value.var="estimate")
fmg_data <- fmg_data[fmg_data$method=="strict",]
fmg_data <- fmg_data[!is.na(fmg_data$`odds ratio`),]

cor(fmg_data$a, fmg_data$`odds ratio`, method = "spearman")
cor.test(fmg_data$a, fmg_data$`odds ratio`, method = "spearman")

fmg_data <- merge(sig_odds, fmg_data, by.x = "GPSC", by.y = "GPSC")
fmg_data$disease <- fmg_data$log_odds>0
mylogit <- glm(disease ~ a, data = fmg_data[fmg_data$issig,], family = "binomial")
summary(mylogit)

mylogit <- glm(disease ~ v, data = fmg_data[fmg_data$issig,], family = "binomial")
summary(mylogit)
```

We can now make a prettier picture that focuses on these results.

```{r}
plotdf <- fmg_params[fmg_params$method=="strict",]
plotdf <- plotdf[plotdf$param %in% c('a','v','G','M','unique serotypes','odds ratio','r/m'),]

library(wesanderson)
library(patchwork)

plotdf$param[plotdf$param=='G'] <- "mean number of genes per genome (G)"
plotdf$param[plotdf$param=='M'] <- "pangeome pool size (M)"
plotdf$param[plotdf$param=='a'] <- "gene gain rate (a)"
plotdf$param[plotdf$param=='v'] <- "gene loss rate (v)"
plotdf$param[plotdf$param=='odds ratio'] <- "log odds ratio"
plotdf$param <- factor(plotdf$param, levels = c("gene gain rate (a)",
                                             "gene loss rate (v)",
                                             "mean number of genes per genome (G)",
                                             "pangeome pool size (M)",
                                             'log odds ratio','r/m','unique serotypes'))
plotdf$estimate[plotdf$param=="log odds ratio"] <- log(plotdf$estimate[plotdf$param=="log odds ratio"])
plotdf$param2 <- as.character(plotdf$param)
plotdf$param2[(plotdf$param=="log odds ratio") & (plotdf$GPSC %in% sig_odds$GPSC[sig_odds$issig])] <- "sig"

cols <- c(wes_palette("Darjeeling1")[c(1,2)],
          wes_palette("Darjeeling2")[c(1)],
          wes_palette("Darjeeling2")[c(2)], 
          wes_palette("Darjeeling1")[c(4)],
          wes_palette("Darjeeling2")[c(4)],
          wes_palette("Darjeeling2")[[3]],
          wes_palette("Darjeeling1")[[3]])

ggplot(plotdf, aes(x=GPSC, y=estimate, fill=param2)) + 
  geom_col() +
  facet_wrap(~ param, ncol = 1, scales = "free_y") +
  theme_bw(base_size = 14) +
  scale_fill_manual(values = cols) +
  theme(legend.position = "none")
```

We can also focus in on the association between gene gain/loss rates and the log odds ratio of invasive disease 

```{r}
tmp <- plotdf[plotdf$param=="gene gain rate (a)" ,]
tmp <- tmp$GPSC[order(-tmp$estimate)]
plotdf$GPSC <- factor(plotdf$GPSC, levels = tmp)

cols <- c(wes_palette("Darjeeling1")[c(1,2)], #k
          wes_palette("Darjeeling2")[c(1)], #k
          wes_palette("Darjeeling2")[[3]])

ggplot(plotdf[plotdf$param %in% c("gene gain rate (a)", "gene loss rate (v)", 'log odds ratio'),], 
       aes(x=GPSC, y=estimate, fill=param2)) + 
  geom_col() +
  facet_wrap(~ param, ncol = 1, scales = "free_y") +
  theme_bw(base_size = 14) +
  scale_fill_manual(values = cols) +
  theme(legend.position = "none")
```

We can also look at the raw Panaroo gene counts

```{r}
pa_files <- Sys.glob("pneumo_gps_data/binary_pa_matrices/*.Rtab")

pa_matrices <- map(pa_files, ~ fread(.x, data.table = FALSE, drop = 1))

raw_sizes_df <- tibble(GPSC=as.numeric(gsub("_s.*", "", gsub(".*pa_matrices/", "", pa_files))),
      method=gsub(".*_s", "s", gsub("_gene.*", "", pa_files)),
       ncore=map_dbl(pa_matrices, ~ sum(rowSums(.x)>(0.98*ncol(.x)))),
       ngenes=map_dbl(pa_matrices, ~ nrow(.x)))

write.csv(raw_sizes_df, file = "pneumo_gps_data/raw_sizes.csv",
          quote = FALSE, row.names = FALSE)
```

###References
